\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-23 NAR_DRAFT.tex

\begin{figure}
  \begin{center}
  \includegraphics{figures/real_signatures_only.pdf}
  \end{center}
  \bkmcomment{tumor in C is probably normal breast plus APOBEC. Mention somewhere?}
  \caption{Tumor mutation profiles.
  In each panel, the x-axis corresponds to each of the 96 possible substitution types, and the y-axis is the proportion of total mutations of each type.
  \textbf{(A)} The observed mutation profile of an acute myeloid leukemia used in our real data analysis \cite{Griffith2015}.
  \textbf{(B)} The observed mutation profile of a breast tumor used in our real data analysis \cite{Shi2018}.
  \textbf{(C)}\&\textbf{(D)} The observed mutation profiles of two breast tumors \cite{Alexandrov2019}.
   }
  \label{NAR-realsigfig}
 \end{figure}

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/sim_signatures_only.pdf}
  \end{center}
  \bkmcomment{This can be smaller. regenerate with its own base\_height}
  \caption{Tumor mutation profiles.
  In each panel, the x-axis corresponds to each of the 96 possible substitution types, and the y-axis is the proportion of total mutations of each type.
  \textbf{(A)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(B)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(C)} Equal proportions of signatures 1, 3, \& 5.
   }
  \label{NAR-simsigfig}
 \end{figure}

@@

\subsection{Real data}
We analyzed two real data sets, one from an acute myeloid leukemia (AML)~\cite{Griffith2015} and one from a multi-region sequencing experiment in breast cancer~\cite{Shi2018}.
We downloaded the normal and primary whole-genome AML tumor bam files from dbGaP accession number phs000159.v8.p4.
Griffith et al.\ generated a platinum set of variant calls for this tumor~\cite{Griffith2015}, which we used for our true positive dataset.
We downloaded the normal and tumor whole-exome breast cancer bam files from NCBI Sequence Read Archive accession SRP070662.
Shi et al.\ generated a gold set of variant calls for each tumor region sequenced~\cite{Shi2018}, which we used for our true positive dataset.
We called variants using Mutect 1.1.7 as in our simulations, except that both these data sets were originally aligned to GRCr37, so we used that reference.

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/kl_only.pdf}
  \end{center}
  \bkmcomment{this would probably look OK shorter too}
  \caption{Convergence of the site-specific prior to the data generating distribution. Plotted is the Kullback-Leibler divergence between the estimated and simulated profiles versus number of incorporated mutations, for \textbf{(A)} whole exomes and \textbf{(B)} whole genomes.
  }
\label{NAR-kl_fig}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=.45\textwidth]{figures/ppv_wes.pdf}
  \end{center}
  \caption{Model performance on 500X simulated whole exomes. 
  Positive predictive value vs. Recall for five different probability cutoffs. 
  Note the restricted axes to improve readability of points}
\label{NAR-ppv_fig}
\end{figure}

>>>>

\begin{figure}
  \begin{center}
  \includegraphics{figures/real_signatures_only.pdf}
  \end{center}
  \bkmcomment{tumor in C is probably normal breast plus APOBEC. Mention somewhere? Fix new typo in panel B.}
  \caption{Tumor mutation profiles.
  In each panel, the x-axis corresponds to each of the 96 possible substitution types, and the y-axis is the proportion of total mutations of each type.
  \textbf{(A)} The observed mutation profile of an acute myeloid leukemia used in our real data analysis \cite{Griffith2015}.
  \textbf{(B)} The observed mutation profile of a breast tumor used in our real data analysis \cite{Shi2018}.
  \textbf{(C)}\&\textbf{(D)} The observed mutation profiles of two breast tumors \cite{Alexandrov2019}.
   }
  \label{NAR-realsigfig}
 \end{figure}

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/sim_signatures_only.pdf}
  \end{center}
  \caption{Tumor mutation profiles.
  In each panel, the x-axis corresponds to each of the 96 possible substitution types, and the y-axis is the proportion of total mutations of each type.
  \textbf{(A)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(B)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(C)} Equal proportions of signatures 1, 3, \& 5.
   }
  \label{NAR-simsigfig}
 \end{figure}

@@

\subsection{Real data}
We analyzed two real data sets, one from an acute myeloid leukemia (AML)~\cite{Griffith2015} and one from a multi-region sequencing experiment in breast cancer~\cite{Shi2018}.
We downloaded the normal and primary whole-genome AML tumor bam files from dbGaP accession number phs000159.v8.p4.
Griffith et al.\ generated a platinum set of variant calls for this tumor~\cite{Griffith2015}, which we used for our true positive dataset.
We downloaded the normal and tumor whole-exome breast cancer bam files from NCBI Sequence Read Archive accession SRP070662.
Shi et al.\ generated a gold set of variant calls for each tumor region sequenced~\cite{Shi2018}, which we used for our true positive dataset.
We called variants using Mutect 1.1.7 as in our simulations, except that both these data sets were originally aligned to GRCr37, so we used that reference.
\rngcomment{Did MuTect fail to call anything we know is there?}

@@

\begin{figure}
  \begin{center}
  \includegraphics{figures/kl_only.pdf}
  \end{center}
  \caption{Convergence of the site-specific prior to the data generating distribution. Plotted is the Kullback-Leibler divergence between the estimated and simulated profiles versus number of incorporated mutations, for \textbf{(A)} whole exomes and \textbf{(B)} whole genomes.
  }
\label{NAR-kl_fig}
\end{figure}

@@
